---
title: "Nate Harrington WHYY"
output: html_document
date: "2025-05-5"
---

<link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/bootstrap@4.3.1/dist/css/bootstrap.min.css" integrity="sha384-ggOyR0iXCbMQv3Xipma34MD+dH/1fQ784/j6cY/iJTQUOhcWr7x9JvoRxT2MZw1T" crossorigin="anonymous">

<div class="container">

<div class="row">

  <div class="col">

<style>
.memohed{
  color: #555555;
  font-size: 32px;
  font-style: italics;
}
p{
  font-size:1.5em;
}

</style>

<h1 class="memohed">Nate Harrington, Second Draft of Pitch</h1>

</div>

</div>

<div class="row">
  <div class="col">

Philadelphia is no exception to the increase in percentage of retirement age citizens, but the city is seeing more of its elderly living below the poverty line than the national average. According to the U.S. Census Bureau, 23% of Philadelphians 65 or older live below the poverty line, 12-points higher than the national average, so how can people with limited income take full advantage of their earned benefits where they apply such as Social Security, SNAP and Medicaid? And how can people approaching the retirement age make sure they can live their best lives once they can no longer work? The current administration is slashing programs and their personnel, making the process harder on everyone, and, with more and more of the nation reaching retirement age, this demand is only going to grow. This story serves people who either are at retirement age, reaching it soon, or are maybe taking care of someone who is off that age so they can best assist this person in receiving benefits.

  </div>
</div>

<div class="row">
  <div class="col-6">
  <iframe title="Percentage of populations living below 100% of the poverty line." aria-label="Grouped Columns" id="datawrapper-chart-YBu2E" src="https://datawrapper.dwcdn.net/YBu2E/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="400" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(a){if(void 0!==a.data["datawrapper-height"]){var e=document.querySelectorAll("iframe");for(var t in a.data["datawrapper-height"])for(var r,i=0;r=e[i];i++)if(r.contentWindow===a.source){var d=a.data["datawrapper-height"][t]+"px";r.style.height=d}}}))}();
</script>  
  </div>
  <div class="col-6">
<p style="font-family:sans-serif; color:#1A936F;">I am going to interview people from organizations like AARP and the Legal Aid Society to get the technical background on how people can access their benefits. I will also look for people who are struggling to access said benefits or might be living below the poverty line. Although there is no guarantee of me finding these people here (I might ask for assistance from people at WHYY for this), I will visit community centers that host events for senior citizens and kitchens that help people get meals to see if I can get an interview. This story is going to be another resource for people struggling with accessing benefits and may help people identify benefits they’ve earned or have access to that they were unaware they had access to. On top of the straightforward benefits of Social Security and other social safety nets, this story (or stories) could help identify strategies for keeping yourself afloat after working age.This story has so many more aspects and so much more relevancy to it than I first initially thought. Each individual program has its own requirements and process to access benefits. Also, an interesting statistic is: there is a higher percentage of retirement age people living under the poverty line than everybody living under the poverty in Philadelphia (23% and 20.3% respectively), a relationship that is flipped nationally and in the larger Philadelphia, Camden, Wilmington metropolitan area. </p>
  </div>
</div>
The most relevant and analogous story to mine is this one about the changes coming to Social Security. My story differs as its more of an explainer about how to gain access to your benefits amidst the cuts, not the differences. This story would include both data visualizations to put an emphasis on the higher rates of poverty in retirement age Philadelphians and some graphics showing some examples of how to follow the steps of applying for social safety net benefits. Photos of the interviewees, of gatherings of retirement age peoples, and more also could be beneficial to the story. Besides the reporting aspects, I would like to take a deeper look at the specific failings of the social safety nets, by finding how much Social Security pays out or how much Medicaid factors in to saving a person money in relation to how much it costs to live in Philadelphia to find how much or little the safety nets helps a person struggling with making ends meet.
	
	Depending on interviews (how easy it is for me to get them), I think I could deliver a story package by late-June to mid-July.









A demo notebook for Dow Jones News Fund data interns

During the week’s training, you will work on a project called Investigate Your Community to produce detailed story pitches relevant to their summer internship positions. The goal is to have the students research a topic relevant to your newsroom so they can get to work immediately at the internship. 

Here are some examples from last year’s DJNF data journalism training:

Hannah Rosenberger, at ASI Media/Counselor magazine, produced this analysis on name, image and likeness trends: https://hrosenberger.github.io/ASI-NIL-Project/

Shifra Dayak, an intern at the Connecticut Mirror, produced this analysis of environment spill data: https://shifradayak.github.io/DJNF_Merrill/ct_env_spills.html.

Michael Zhang, an intern at the Houston Landing, explored local election data: https://michaelzhangreporter.github.io/Houston_Campaign_Finance/


Here's what your newsroom suggested for a project
Here are some newsroom-wide projects coming up that we welcome data journalism contributions from our future DJ interns.
 
Aging in the region – tentative project title is “Growing Golden: Aging with Purpose.” Numerous stories about aging and thriving in retirement will cover a range of financial, health, housing, dying, family coping and planning for retirement story topics. This is an explanatory journalism project.
 
Developing project on Black Women in the region – inspired by the election of Philadelphia’s first woman mayor and Black woman mayor. Areas of storytelling will include: politics, education and health
 
Philadelphia neighborhood guides is a project coming from Billy Penn.
 
An existing project on education is such, and we will continue to produce stories through mid-summer -- Creativity Sparks Success Project https://whyy.org/series/creativity-sparks-success/



This notebook helps you get started on the Growing Golden project.


# Load software libraries
```{r}
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("readxl")
#install.packages("rvest")
#install.packages("jsonlite")
library(tidyverse)
library(janitor)
library(readxl)
library(rvest)
library(jsonlite)

```


# Load your data

Check Census Reporter for ideas. Here's a start
https://censusreporter.org/data/table/?table=B01001&primary_geo_id=16000US4260000&geo_ids=16000US4260000,05000US42101,31000US37980,04000US42,01000US


```{r}

# 
# YOUR_DATA <- read.csv("../data/XXXXX.csv") |> 
#     clean_names()

# data <- read_excel("../data/2024-12-Revenue.xlsx", skip=2) |> 
#     clean_names()

data <- read.csv("../data/acs2023_1yr_B01001_04000US42.csv") |> 
    clean_names() 
```

Restructure the table for 

```{r}


demographic_table <- data %>%
  # First select just the relevant columns
  select(name, matches("^b01001\\d+$")) %>%  
  # Pivot to make each b01001 variable a row
  pivot_longer(cols = starts_with("b01001"),
               names_to = "variable",
               values_to = "value") %>%
  # Then pivot to make each location a column
  pivot_wider(names_from = name,
              values_from = value) %>%
  # Clean up column names if needed
  rename_with(~ gsub(", PA$", "", .x), matches("Philadelphia County")) %>%
  rename_with(~ gsub(" Metro Area$", "", .x), matches("Philadelphia-Camden")) |> 
  clean_names()

```


Read json file as a dataframe
```{r}

data_labels <- fromJSON("../data/metadata.json")

column_list <- data_labels[["tables"]][["B01001"]][["columns"]]

# Create a dataframe from the list
label_df <- data.frame(
  variable = names(column_list),
  indent = sapply(column_list, function(x) x$indent),
  name = sapply(column_list, function(x) x$name),
  stringsAsFactors = FALSE
) |> 
  mutate(variable = tolower(variable))


```


# Join labels and data
```{r}
demographic_table1 <- demographic_table |> 
  inner_join(label_df, by=c("variable")) |> 
  select(name, philadelphia_pa,philadelphia_county, pennsylvania,  
          philadelphia_camden_wilmington_pa_nj_de_md,  united_states, variable, indent)

#fix to specify male-female
demographic_table1 <- demographic_table1 %>%
  mutate(name = case_when(
    # Skip rows 1, 2, 26, and 50+ (if any)
    row_number() %in% c(1, 2, 26) ~ name,
    # Add " (male)" to rows 3-25
    row_number() >= 3 & row_number() <= 25 ~ paste0(name, "_male"),
    # Add " (female)" to rows 27-49
    row_number() >= 27 & row_number() <= 49 ~ paste0(name, "_female"),
    # Keep any other rows unchanged
    TRUE ~ name
  ))

```


Data details:


# Explore the data types
```{r}

#glimpse(YOUR_DATA)

glimpse(demographic_table1 )

```


# Questions 

Write down questions you want answered:



# Data Details
https://censusreporter.org/data/table/?table=B01001&primary_geo_id=16000US4260000&geo_ids=16000US4260000,05000US42101,31000US37980,04000US42,01000US


Next steps:
You should learn to use the TidyCensus package in R. It would make this work so much easier.
https://walker-data.com/tidycensus/


# Second draft of pitch



</div>